F1000 paper figures

### 2017-06-20 inst/paper/figs.Rmd
fixed quotes and a few minor changes

## Figure 2
```{r fig2}
# `valr_example()` identifies the path of example files
bedfile <- valr_example('genes.hg19.chr22.bed.gz')
genomefile <- valr_example('hg19.chrom.sizes.gz')
bgfile  <- valr_example('hela.h3k4.chip.bg.gz')

@@

p <- ggplot(res, aes(x = .win_id, y = win_mean)) +
  geom_point(size = 0.25) + geom_pointrange(sd_limits, size = 0.1) + 
  scale_x_continuous(labels = x_labels, breaks = x_breaks) + 
  xlab('Position (bp from TSS)') + ylab('Signal') + 
  theme_classic() +
  pub_theme

plot_grid(p, align = 'h', nrow = 1, labels="AUTO")
ggsave("figure2.pdf", width = 5.6, height = 4)
```

@@

# Set-up test genome
genome <- read_genome(valr_example('hg19.chrom.sizes.gz'))
write_tsv(genome, "genome.txt", col_names=FALSE)

@@

```{r}
library(valr)
# function to retrieve path to example data
bed_filepath <- valr_example("3fields.bed.gz") 
read_bed(bed_filepath)

@@

Using mysql
```{r, warnings = F}
library(dplyr, warn.conflicts = F)
hg19 <- db_ucsc("hg19")
tbl(hg19, "refGene")
```

@@

repeats <- read_bed(valr_example('hg19.rmsk.chr22.bed.gz'), n_fields = 6) 
genome <- read_genome(valr_example('hg19.chrom.sizes.gz'))
genes <- read_bed12(valr_example('hg19.refGene.chr22.bed.gz'))
# convert bed12 to bed with exons
exons <- bed12_to_exons(genes)

# function to repeat interval shuffling
shuffle_intervals <- function(n, .data, genome) {
  replicate(n, bed_shuffle(.data, genome), simplify = FALSE) %>%
    bind_rows(.id = 'rep') %>%
    group_by(rep) %>% nest()
}
nreps <- 100
shuffled <- shuffle_intervals(n = nreps, repeats, genome) %>%
  mutate(jaccard = data %>%
           map(bed_jaccard, repeats) %>%
           map_dbl("jaccard"))
shuffled

>>>>

## Figure 2
```{r fig2}
bedfile <- valr_example("genes.hg19.chr22.bed.gz")
genomefile <- valr_example("hg19.chrom.sizes.gz")
bgfile  <- valr_example("hela.h3k4.chip.bg.gz")

@@

p <- ggplot(res, aes(x = .win_id, y = win_mean)) +
  geom_point(size = 0.25) + geom_pointrange(sd_limits, size = 0.1) + 
  scale_x_continuous(labels = x_labels, breaks = x_breaks) + 
  xlab("Position (bp from TSS)") + ylab("Signal") + 
  theme_classic() +
  pub_theme +
  theme(axis.text.x = element_text(angle = 90))

plot_grid(p, align = "h", nrow = 1, labels="AUTO")
ggsave("figure2.pdf", width = 5.6, height = 5.6)
```

@@

# Set-up test genome
genome <- read_genome(valr_example("hg19.chrom.sizes.gz"))
write_tsv(genome, "genome.txt", col_names=FALSE)

@@

### file I/O
```{r}
library(valr)
# function to retrieve path to example data
bed_filepath <- valr_example("3fields.bed.gz") 
read_bed(bed_filepath)

@@

### syntax demo

```{r syntax_demo, message = FALSE}
library(valr)
library(dplyr)

snps <- read_bed(valr_example("hg19.snps147.chr22.bed.gz"), n_fields = 6)
genes <- read_bed(valr_example("genes.hg19.chr22.bed.gz"), n_fields = 6)

# find snps in intergenic regions
intergenic <- bed_subtract(snps, genes)
# distance from intergenic snps to nearest gene
nearby <- bed_closest(intergenic, genes)

nearby %>%
  select(starts_with("name"), .overlap, .dist) %>%
  filter(abs(.dist) < 1000)
```

### Figure 1 code demo

```{r}
x <- tibble::tribble(
  ~chrom, ~start, ~end,
  "chr1", 25,     50,
  "chr1", 100,    125
)

y <- tibble::tribble(
  ~chrom, ~start, ~end,
  "chr1", 30,     75
)

bed_glyph(bed_intersect(x, y))
```

```{r}
x <- tibble::tribble(
  ~chrom, ~start, ~end,
  "chr1",      1,      50,
  "chr1",      10,     75,
  "chr1",      100,    120
)

bed_glyph(bed_merge(x))
```

### grouping data demo

```{r}
x <- tibble::tribble(
  ~chrom, ~start, ~end, ~strand,
  "chr1", 1,      100,  "+",
  "chr1", 50,     150,  "+",
  "chr2", 100,    200,  "-"
)

y <- tibble::tribble(
  ~chrom, ~start, ~end, ~strand,
  "chr1", 50,     125,  "+",
  "chr1", 50,     150,  "-",
  "chr2", 50,     150,  "+"
)

# intersect tbls by strand
x <- group_by(x, strand)
y <- group_by(y, strand)

bed_intersect(x, y)
```

Comparisons between intervals on opposite strands are done using the `flip_strands()` function:

```{r}
x <- group_by(x, strand)

y <- flip_strands(y)
y <- group_by(y, strand)

bed_intersect(x, y)
```

### Figure 2 code demo
```{r }
bedfile <- valr_example("genes.hg19.chr22.bed.gz")
genomefile <- valr_example("hg19.chrom.sizes.gz")
bgfile  <- valr_example("hela.h3k4.chip.bg.gz")

genes <- read_bed(bedfile, n_fields = 6)
genome <- read_genome(genomefile)
y <- read_bedgraph(bgfile)
```

Then we generate 1 bp intervals to represent transcription start sites (TSSs). We focus on `+` strand genes, but `-` genes are easily accomodated by filtering them and using `bed_makewindows()` with `reversed` window numbers.

```{r}
# generate 1 bp TSS intervals, "+" strand only
tss <- genes %>%
  filter(strand == "+") %>%
  mutate(end = start + 1)

x
```

Now we use the `.win_id` group with `bed_map()` to caluclate a sum by mapping `y` signals onto the intervals in `x`. These data are regrouped by `.win_id` and a summary with `mean` and `sd` values is calculated.

```{r}
# map signals to TSS regions and calculate summary statistics.
res <- bed_map(x, y, win_sum = sum(value, na.rm = TRUE)) %>%
  group_by(.win_id) %>%
  summarize(win_mean = mean(win_sum, na.rm = TRUE),
            win_sd = sd(win_sum, na.rm = TRUE))

res
```

Finally, these summary statistics are used to construct a plot that illustrates histone density surrounding TSSs.

```{r}
library(ggplot2)

p <- ggplot(res, aes(x = .win_id, y = win_mean)) +
  geom_point(size = 0.25) + geom_pointrange(sd_limits, size = 0.1) + 
  scale_x_continuous(labels = x_labels, breaks = x_breaks) + 
  xlab("Position (bp from TSS)") + ylab("Signal") + 
  theme_classic()
```

@@

repeats <- read_bed(valr_example("hg19.rmsk.chr22.bed.gz"), n_fields = 6) 
genome <- read_genome(valr_example("hg19.chrom.sizes.gz"))
genes <- read_bed12(valr_example("hg19.refGene.chr22.bed.gz"))
# convert bed12 to bed with exons
exons <- bed12_to_exons(genes)

# function to repeat interval shuffling
shuffle_intervals <- function(n, .data, genome) {
  replicate(n, bed_shuffle(.data, genome), simplify = FALSE) %>%
    bind_rows(.id = "rep") %>%
    group_by(rep) %>% nest()
}
nreps <- 100
shuffled <- shuffle_intervals(n = nreps, repeats, genome) %>%
  mutate(jaccard = data %>%
           map(bed_jaccard, repeats) %>%
           map_dbl("jaccard"))
shuffled
